---
title: Project title
author: by Mr. Palmer's Penguins
output: github_document
---

```{r load-packages, echo = FALSE}
# load packages here
library(lubridate)
library(countrycode)
library(gganimate)
library(plyr)
library(dplyr)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(showtext)) # cool fonts
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggtext)) # superscript in plot labels
```

```{r load-data, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))

```

```{r set-options, echo = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.asp= 0.618, out.width = "80%")

```

## Introduction

The set, which includes `r ncol(olympics)` variables and `r nrow(olympics)` observations, provides insight into the athletes that competed in and the results of the Olympic games from Athens 1896 to Rio 2016. Each observation includes information about an athlete on a per event basis. Therefore, if an athlete competes in multiple different events, there will be multiple different observations reflecting each event an individual participated in. Each observation details medal results (Gold, Silver, Bronze, or NA), the athlete's background (age, sex, height, weight, team they're competing for, etc.), and context about the event (where and when it was held, season, sport the athlete competes in, etc.).

We chose this dataset because we think it can provide a lot of good insights into some of the trends at the Olympics over the past century as well as provide strong opportunities for our group to develop our visualization capabilities. 

This dataset was accessed on [kaggle.com](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results) and includes data scraped in May 2018 from Sports Reference/OlympStats, sports statistics sites, by Randi Griffin, a Senior Data Scientist at Boston Consulting Group. Moreover, it was selected as part of the [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021#readme7c70d95441aec295a1e92da9d71e2872877d663c) challenge on 7/27/21.

## Do certain physical characteristics (such as body mass index, age, and sex) of Olympic participants differ by sport and/or change with time?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

As we know, the Olympic Games is a unique opportunity for diverse and talented individuals to showcase their skills across a variety of disciplines. As a result, we are interested in better understanding how the diversity of physical characteristics in the set of competing athletes potentially differs by sport and/or has changed with time. Three physcial characteristics of interest are body mass index (a ratio of `height` and `weight`), `sex`, and `age`. It's often the case that the media/press discuss these these physical characteristics as key metrics when describing athletes. Therefore, we are curious about better understanding the extent to which these differ (if at all).

Firstly, we are hoping to understand if the optimal build of an Olympic athlete differs by `sex` or by `sport`. For example, a high performing male Olympic wrestler is likely to have  a different build versus a typical female gymnast. Pretty disparate builds, right? Even within one sport, say gymnastics, we hypothesize that there may be different optimal heights and weights for successful athletes of different sexes.

Secondly, we will assess if the mean age of Olympic participants for a given `year` has changed with time by `sex`. We know that over time the Olympics have transitioned from an amateur to professional competition. Therefore, we are curious to see if this has any association with the ages of participants since we know that certain age are more optimal for "peak performance" than others.
We will separate the visualization by `sex` because while there an approximately equal number of males and females in the world, we know that access to athletics participation is not always equal which could be associated with participation rates by `age`.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.


For our first plot, we plan to use boxplots to plot the distribution of a size characteristic by `sex` (since characteristics tend to differ among males and females) and faceting by `sport`. We are unsure which of the `height`, `weight`, `height / weight`, or `bmi` ratio will be most telling when examined across sexes and sports. We will start by making 4 plots (each with one of these on the y-axis) and then select the plot where the trends are most insightful.

Two of these variables require new calculations in the data wrangling step. In order to plot the height-weight-ratio, we will have to create two new calculated columns. First, a `height_weight_ratio` column that simply divides an athlete's `height` by their `weight`, and a more complicated `BMI` column that represents an existing metric--[BMI](https://en.wikipedia.org/wiki/Body_mass_index)--that integrates height and weight into a single measure. . 

Based on the results from these preliminary plots, in addition to narrowing in on a size characteristic, we might pick a few select sports to focus in on (i.e. most popular sports) or group the sports based on common characteristics (ex: contact vs. non-contact, etc.) since there are 66 sports in the dataset.

For our second plot, we are going to make a scatter plot of average `age` vs. `year` color mapped by `sex`. To start, we will find the average age of the participants by sex for each year. This will allow us to a not too crowded scatter plot since there are now only two observations (one for Male and one for Female) for each year. We choose a scatter plot because we believe it can best best represent the relationship and highlight potential trends. We considered a line plot but ultimately decided that the autonomy of each of the points (rather than them being connected) would better allow us to compare observations `year` by `year`. Since there is. unique context that surrounds each Olympics game, we feel as though a more year-specific approach could be more meaningful.

### Analysis

``` {r wrangling-plot-one}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(height_weight_ratio = height / weight,
         BMI = 10000 * weight / (height * height))

# load Olympic font
font_add_google(name = "Oswald")
showtext_auto()
```

After examining each of `height`, `weight`, `height_weight_ratio`, and `BMI`, we have chosen to proceed with `BMI` as our size characteristic for this plot, as the visualizations were the most accessible and intuitive to interpret. After examining the plot displaying the `BMI` for all sports, we are choosing a few select groups of sports to include. The next step is to create separate dataframes for each of these categories. We also want to represent each athlete once on the plot, so we will have to use `distinct()`.

``` {r mutate-category}
# create variable to represent category of sport
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  filter(sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling", "Gymnastics", "Trampolining", "Archery", "Shooting", "Athletics", "Rowing")) %>%
  mutate(category = case_when(
                    sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling") ~ "weightclass",
                    sport %in% c("Gymnastics", "Trampolining") ~ "acrobatic",
                    sport %in% c("Archery", "Shooting") ~ "coordination",
                    TRUE ~ "diverse"),
        category = factor(category, levels = c("acrobatic", "diverse", "coordination", "weightclass")))

# create separate dataframes for each of these categories
olympics_weightclass <- olympics %>%
  filter(category == "weightclass")

olympics_coordination <- olympics %>%
  filter(category == "coordination")

olympics_diverse <- olympics %>%
  filter(category == "diverse")

olympics_acrobatic <- olympics %>%
  filter(category == "acrobatic")
```

We are now ready to create a separate plot for each sport category. We will also silence messages which indicate that rows with NA have been dropped from the dataframe when creating the boxplots.

``` {r plot-weightclass, warning = FALSE}
ggplot(olympics_weightclass, mapping = aes(y = fct_rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in weightclass-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-coordination, warning = FALSE}
ggplot(olympics_coordination, mapping = aes(y = rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in coordination-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-diverse, warning = FALSE}
ggplot(olympics_diverse, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in Olympic sports which encompass many skills",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-acrobatic, warning = FALSE}
ggplot(olympics_acrobatic, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in acrobatic Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

We found a citation for the Tokyo 2020 Olympics logo font [here](https://www.reddit.com/r/identifythisfont/comments/4ig8ua/font_used_on_the_tokyo_2020_logo/) and were pointed to an open-source alternative [here](https://graphicdesign.stackexchange.com/questions/7178/is-there-a-din-font-free-alternative). We got the hex code used in plot text manually from that logo source. We chose the hex codes for our favorite gender color mapping from Telegraph 2018 [here](https://blog.datawrapper.de/gendercolor/).

We can now visualize these groups on the same plot, to get a better sense of broader variability in size characteristics among an array of Olympic sports.

``` {r fonts-and-plot-one, warning = FALSE}
ggplot(olympics, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
  
```

For our second plot, we will look at how `age` of Olympic participants by `sex` has changed with time.

We will re-load the CSV file to clear any manipulations done on the Olympics data frame to make the first plot.

```{r load-data-2, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
```

Since the observations in the data are done on a per event competed in basis (rather than a per athlete basis), we will use `distinct()` to filter the data so we only have unique athletes represented each `year`. In other words, each `year` will contain a set of observations representing one athlete each. This will remove cases where the age of an athlete who competed in multiple events for a given Olympics is repeated multiple times.

```{r distinct-year}
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, year, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")))
```

Finally, we will make the plot showing the average yearly `age` of athletes by `sex`. We will drop observations in which the `age` is unknown.

```{r age-plot}
olympics %>%
  drop_na(age) %>%
  group_by(year, sex) %>%
  summarise(avg_age = mean(age)) %>%
  arrange(desc(sex)) %>%
  ggplot(mapping = aes(y = avg_age, x = year)) +
  geom_point(aes(color = sex)) +
  labs(x = "Year",
       y = "Age",
       color = "Sex",
       title = "Average Yearly Age of Athletes By Sex",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB"), guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

The mean was used, rather than the median, of `age` as this better captures the range of ages of athletes for a given year.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

In the "Sex vs Body Mass Index Distribution" plots, men typically have a higher-centered distribution of `BMI` than women in the same sport, but other trends are more apparent by `category`. In the first plot, which includes wrestling, weightlifting, judo, and boxing, we see a strong right skew in most distributions. We interpret this as a reflection of the weight classes present in those sports, with many participants entering in lower weight classes, giving a lower center, but a non-insignificant number of athletes with BMIs higher than 30. However, there are fewer at these high BMIs, possibly due to the difficulty in maintaining athletic competitiveness at that proportion. Shooting and archery also have generally higher BMIs than other sports, and again a right-skew is observed in the distributions. This reflects those sports' focus on hand-eye coordination, and thus athlete body type or height or weight are more variable.

The next group of sports, Athletics and Rowing, feature distributions with large variability on both sides. We interpret this as illustrating the diversity of body types considered desirable in this sports: from slender, lithe coxswains and sprinters to heavier, more powerful rowers and throwers. Finally, the largest disparities between `sex` in a particular sport are observed in Gymnastics and Trampoline. Here, we speculate that the female gymnastics disciplines favor petite, nimble athletes, as opposed to the more upper-body strength oriented male apparatuses, which contribute to a higher average BMI for male Olympic gymnasts.

Finally, for the "Average Yearly Age of Athletes By Sex" plot, it was found that the mean `age` of athletes by `sex` has changed pretty significantly over the years. For instance, the range of mean ages participating has converged pretty dramatically. Prior to the 1970s, the mean ages fluctuated in the 20-30 years old range with even some significant outliers hitting 35 and 50 years old in some cases. However, after the 1970s, it appears as though the mean ages converged in the 25-27 years old range. This makes sense because, according to a Britannica article found [here](https://www.britannica.com/sports/Olympic-Games), the 1970s were a turning point for the games moving from allowing professional athletes to compete rather than just amateurs. The article also notes that it wasn't until the 1980s that fully-fledged professional participation was permitted. With this information, we can speculate that there might be an association between the allowing of professional athletes and the convergence. Perhaps since most professional athletes are around 25 years old (a point of peak physical ability of many), the mean age reflects the demographic shift (amateur to professional) of participants. 

In addition to age converging, it is interesting to note the difference between Males and Females. For all years but 3, the mean age of males was higher than females. Perhaps since males take longer to physically mature (puberty is at a later age), that could explain why their ability to compete at "peak perform" in a professional context is at a later age. Moreover, it's possible that the extreme variability between mean ages of females prior to 1950 is the result of fewer females competing. With fewer competitors, it's more likely that outlier events (extremely old participants) can skew the mean. For example, prior to 1920, there were 210 female participants versus 9,790 male participants as depicted below.

```{r}
olympics %>%
  filter(year <= 1920) %>%
  count(sex)
```

These numbers are the total number of unique participants participating in each Olympics game prior to 1920.

## Question 2 <- Update title to relate to the question you’re answering

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.
#### Plot 1

```{r}
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
noc_regions <- read_csv(file = paste0(here::here(), "/data/noc_regions.csv"))
```

```{r wrangling-medals}
# make appropriate date
olympics %<>%
  mutate(date = ifelse(season == "Winter", paste0(year, "-02-01"), paste0(year, "-07-01")),
         date = ymd(date),
         medal_winner = ifelse(is.na(medal), 0, 1),
         medal_score = case_when(
           medal == "Bronze" ~ 1,
           medal == "Silver" ~ 2,
           medal == "Gold" ~ 3,
           TRUE ~ 0))

noc_regions %<>%
  janitor::clean_names() %>%
  mutate(region = ifelse(region == "Boliva", "Bolivia", region))

flags <- tibble(noc = noc_regions$noc, region = noc_regions$region, flag = countrycode(noc_regions$region, 'country.name', 'unicode.symbol'))

olympics <- left_join(olympics, flags, by = "noc")

olympics
```

```{r}
olympics_filter <- olympics %>%
  filter(season == "Summer") %>%
  group_by(noc) %>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score))%>%
  filter(total_winners >= 910)

```



```{r}
olympics_filter

```
```{r}

ggplot(olympics_filter, aes(x = reorder(noc, total_winners), y = total_winners, fill = noc, color = noc))+
  geom_bar(stat='identity')+
  labs(x = "Country",
       y = "Medal Count",
       title = "All-time Summer Olympic Medal Counts",
       subtitle = "For the top 10 most winningest countries in history",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  scale_color_manual(values = c("#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB")) +
  scale_fill_manual(values = c("#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB")) +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.position = "none")


```

```{r}

usa_fil <- olympics %>%
  filter(noc == "USA") %>%
  group_by(sport, year)%>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score))%>%
  filter(sport %in% c("Swimming", "Athletics"))

usa_fil
```

```{r}

ggplot(usa_fil, aes(x = year, y = total_winners, color = sport, group = sport))+
  geom_line()+
  geom_point()+
  labs(x = "Year",
       y = "Medal Count",
       title = "USA Olympic Medals Won in Athletics and Swimming",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  scale_color_manual(values = c("#7B38EC","#5CC0AB"))+
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.18),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))

```

```{r}

usa_fil2 <- ddply(usa_fil, .(sport), transform, cumsumwin = cumsum(total_winners))

usa_fil2

ggplot(usa_fil2, aes(x = year, y = cumsumwin, color = sport, group = sport))+
  geom_line()+
  labs(x = "Year",
       y = "Cumulative Medal Count",
       title = "USA Cumulative Olympic Medals Won in Athletics and Swimming",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  scale_color_manual(values = c("#7B38EC","#5CC0AB"))+
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.18),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))

```

In our first visualization, we will plot `medals_total` on the y-axis and `date` on the x-axis with lines grouped by `country` and faceting by `sport`. We will select just a few sports of interest to compare and use selective color mapping for countries of interest and transparency to strongly highlight a few countries and make the remaining lines indistinguishable in the background. If this technique remains visually overwhelming after prototyping our plot, we will filter to select just a few countries of interest in each sport and color map only those.

#### Plot 2

```{r}
gdp <- read_csv("data/gdp-per-capita-maddison-2020.csv")

gdp %<>%
  janitor::clean_names()
```

```{r}
olympics_gdp <- olympics %>%
  group_by(region, date) %>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop") %>%
  rename(entity = region) %>%
  mutate(year = year(date)) %>%
  left_join(gdp, by = c("entity", "year"))
```

```{r}
olympics_gdp %>%
  filter(year >= 2010) %>%
  ggplot(aes(x = gdp_per_capita, total_score)) +
  geom_point() +
  facet_wrap(~date)
```

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
