---
title: Project title
author: by Mr. Palmer's Penguins
output: github_document
---

```{r load-packages, echo = FALSE}
# load packages here
library(lubridate)
library(countrycode)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(showtext)) # cool fonts
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggtext)) # superscript in plot labels
```

```{r load-data, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))

```

```{r set-options, echo = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.asp= 0.618, out.width = "80%")

```

## Introduction

The set, which includes `r ncol(olympics)` variables and `r nrow(olympics)` observations, provides insight into the athletes that competed in and the results of the Olympic games from Athens 1896 to Rio 2016. Each observation includes information about an athlete on a per event basis. Therefore, if an athlete competes in multiple different events, there will be multiple different observations reflecting each event an individual participated in. Each observation details medal results (Gold, Silver, Bronze, or NA), the athlete's background (age, sex, height, weight, team they're competing for, etc.), and context about the event (where and when it was held, season, sport the athlete competes in, etc.).

We chose this dataset because we think it can provide a lot of good insights into some of the trends at the Olympics over the past century as well as provide strong opportunities for our group to develop our visualization capabilities. 

This dataset was accessed on [kaggle.com](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results) and includes data scraped in May 2018 from Sports Reference/OlympStats, sports statistics sites, by Randi Griffin, a Senior Data Scientist at Boston Consulting Group. Moreover, it was selected as part of the [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021#readme7c70d95441aec295a1e92da9d71e2872877d663c) challenge on 7/27/21.

## Do certain physical characteristics (such as body mass index, age, and sex) of Olympic participants differ by sport and/or change with time?

Does the optimal build of an Olympic athlete differ by sex or by sport? 
How have the ages of these Olympic athletes evolved over time?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

We are interested in exploring the physical characteristics of Olympic athletes. The Olympic Games is unique in showcasing a variety of disciplines and the body types of athletes who succeed at them. For example, imagine a typical male Olympic wrestler, versus a typical female gymnast. Pretty disparate builds, right? Even within one sport, say gymnastics, we hypothesize that there may be different optimal heights and weights for successful athletes of different genders  Therefore, to answer our first question, we will visualize the variables `height`, `weight`, `age`, `sex`, `sport`, and `year`.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.


For our first plot, we plan to use boxplots to plot the distribution of a size characteristic by `sex` (since characteristics tend to differ among males and females) and faceting by `sport`. We are unsure which of the `height`, `weight`, `height / weight`, or `bmi` ratio will be most telling when examined across sexes and sports. We will start by making 4 plots (each with one of these on the y-axis) and then select the plot where the trends are most insightful.

Two of these variables require new calculations in the data wrangling step. In order to plot the height-weight-ratio, we will have to create two new calculated columns. First, a `height_weight_ratio` column that simply divides an athlete's `height` by their `weight`, and a more complicated `BMI` column that represents an existing metric--[BMI](https://en.wikipedia.org/wiki/Body_mass_index)--that integrates height and weight into a single measure. . 

Based on the results from these preliminary plots, in addition to narrowing in on a size characteristic, we might pick a few select sports to focus in on (i.e. most popular sports) or group the sports based on common characteristics (ex: contact vs. non-contact, etc.) since there are 66 sports in the dataset.

### Analysis

``` {r wrangling-plot-one}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(height_weight_ratio = height / weight,
         BMI = 10000 * weight / (height * height))

# load Olympic font
font_add_google(name = "Oswald")
showtext_auto()
```

After examining each of `height`, `weight`, `height_weight_ratio`, and `BMI`, we have chosen to proceed with `BMI` as our size characteristic for this plot, as the visualizations were the most accessible and intuitive to interpret. After examining the plot displaying the `BMI` for all sports, we are choosing a few select groups of sports to include. The next step is to create separate dataframes for each of these categories. We also want to represent each athlete once on the plot, so we will have to use `distinct()`.

``` {r mutate-category}
# create variable to represent category of sport
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  filter(sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling", "Gymnastics", "Trampolining", "Archery", "Shooting", "Athletics", "Rowing")) %>%
  mutate(category = case_when(
                    sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling") ~ "weightclass",
                    sport %in% c("Gymnastics", "Trampolining") ~ "acrobatic",
                    sport %in% c("Archery", "Shooting") ~ "coordination",
                    TRUE ~ "diverse"),
        category = factor(category, levels = c("acrobatic", "diverse", "coordination", "weightclass")))

# create separate dataframes for each of these categories
olympics_weightclass <- olympics %>%
  filter(category == "weightclass")

olympics_coordination <- olympics %>%
  filter(category == "coordination")

olympics_diverse <- olympics %>%
  filter(category == "diverse")

olympics_acrobatic <- olympics %>%
  filter(category == "acrobatic")
```

We are now ready to create a separate plot for each sport category. We will also silence messages which indicate that rows with NA have been dropped from the dataframe when creating the boxplots.

``` {r plot-weightclass, warning = FALSE}
ggplot(olympics_weightclass, mapping = aes(y = fct_rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in weightclass-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-coordination, warning = FALSE}
ggplot(olympics_coordination, mapping = aes(y = rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in coordination-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-diverse, warning = FALSE}
ggplot(olympics_diverse, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in Olympic sports which encompass many skills",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-acrobatic, warning = FALSE}
ggplot(olympics_acrobatic, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in acrobatic Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

We found a citation for the Tokyo 2020 Olympics logo font [here](https://www.reddit.com/r/identifythisfont/comments/4ig8ua/font_used_on_the_tokyo_2020_logo/) and were pointed to an open-source alternative [here](https://graphicdesign.stackexchange.com/questions/7178/is-there-a-din-font-free-alternative). We got the hex code used in plot text manually from that logo source. We chose the hex codes for our favorite gender color mapping from Telegraph 2018 [here](https://blog.datawrapper.de/gendercolor/).

We can now visualize these groups on the same plot, to get a better sense of broader variability in size characteristics among an array of Olympic sports.

``` {r fonts-and-plot-one, warning = FALSE}
ggplot(olympics, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
  
```

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

Plot 2 (Drew):

```{r load-data-2, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
```

```{r BMI-create}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(height_weight_ratio = height / weight,
         BMI = 10000 * weight / (height * height))
```

```{r distinct-year}
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, year, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")))
```

```{r age-plot}
olympics %>%
  drop_na(age) %>%
  group_by(year, sex) %>%
  summarise(avg_age = mean(age)) %>%
  arrange(desc(sex)) %>%
  ggplot(mapping = aes(y = avg_age, x = year)) +
  geom_point(aes(color = sex)) +
  labs(x = "Year",
       y = "Age",
       color = "Sex",
       title = "Average Yearly Age of Athletes By Sex",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
  
```
Tried median to see if it gets rid of outlier but does not



For our second plot, we will look more closely at how size characteristics have changed over time, using `year` to make a time series line plots for these different values. In other words, we will plot one of the size characteristics (`height`, `weight`, or the height:weight ratio) on the y-axis vs. time on the x-axis. Lines by `sex` will connect the observations. We might also facet these plots by `age` to see if, for example, height of athletes has changed more or less over time for younger athletes vs. older athletes. In order to facet by `age`, we will have to make `age` a categorical variable by setting different cutoffs in which we group observations. 

The ages in the dataset range from 10 to 97 years old. Therefore, we will plot a histogram and do summary statistics to understand which cutoffs make the most sense given the data. Since athletes tend to be in their mid 20s, we'll likely have smaller differences in ages in the 20-30 age range whereas all athletes 40-97, for example, might be grouped together. However, as mentioned, this would depend on the distribution of the data.

Like in Plot 1, we will evaluate which of `height`, `weight`, or the `height:weight` ratio provides the most insightful visualization. Ideally, the variable we choose in Plot 1 will be the same as in Plot 2; however, we understand creating visualizations is an iterative process are open to evaluating all options as we plot.

For our second plot, we will look more closely at how size characteristics have changed over time, using `year` to make a time series line plots for these different values. In other words, we will plot one of the size characteristics (`height`, `weight`, or the height:weight ratio) on the y-axis vs. time on the x-axis. Lines by `sex` will connect the observations. We might also facet these plots by `age` to see if, for example, height of athletes has changed more or less over time for younger athletes vs. older athletes. In order to facet by `age`, we will have to make `age` a categorical variable by setting different cutoffs in which we group observations. 

The ages in the dataset range from 10 to 97 years old. Therefore, we will plot a histogram and do summary statistics to understand which cutoffs make the most sense given the data. Since athletes tend to be in their mid 20s, we'll likely have smaller differences in ages in the 20-30 age range whereas all athletes 40-97, for example, might be grouped together. However, as mentioned, this would depend on the distribution of the data.

Like in Plot 1, we will evaluate which of `height`, `weight`, or the `height:weight` ratio provides the most insightful visualization. Ideally, the variable we choose in Plot 1 will be the same as in Plot 2; however, we understand creating visualizations is an iterative process are open to evaluating all options as we plot.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

In all plots, men typically have a higher-centered distribution of `BMI` than women in the same sport, but other trends are more apparent by `category`. In the first plot, which includes wrestling, weightlifting, judo, and boxing, we see a strong right skew in most distributions. We interpret this as a reflection of the weight classes present in those sports, with many participants entering in lower weight classes, giving a lower center, but a non-insignificant number of athletes with BMIs higher than 30. However, there are fewer at these high BMIs, possibly due to the difficulty in maintaining athletic competitiveness at that proportion. Shooting and archery also have generally higher BMIs than other sports, and again a right-skew is observed in the distributions. This reflects those sports' focus on hand-eye coordination, and thus athlete body type or height or weight are more variable.

The next group of sports, Athletics and Rowing, feature distributions with large variability on both sides. We interpret this as illustrating the diversity of body types considered desirable in this sports: from slender, lithe coxswains and sprinters to heavier, more powerful rowers and throwers. Finally, the largest disparities between `sex` in a particular sport are observed in Gymnastics and Trampoline. Here, we speculate that the female gymnastics disciplines favor petite, nimble athletes, as opposed to the more upper-body strength oriented male apparatuses, which contribute to a higher average BMI for male Olympic gymnasts.

## Question 2 <- Update title to relate to the question you’re answering

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.
#### Plot 1

```{r}
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
noc_regions <- read_csv(file = paste0(here::here(), "/data/noc_regions.csv"))
```

```{r wrangling-medals}
# make appropriate date
olympics %<>%
  mutate(date = ifelse(season == "Winter", paste0(year, "-02-01"), paste0(year, "-07-01")),
         date = ymd(date),
         medal_winner = ifelse(is.na(medal), 0, 1),
         medal_score = case_when(
           medal == "Bronze" ~ 1,
           medal == "Silver" ~ 2,
           medal == "Gold" ~ 3,
           TRUE ~ 0))

noc_regions %<>%
  janitor::clean_names() %>%
  mutate(region = ifelse(region == "Boliva", "Bolivia", region))

flags <- tibble(noc = noc_regions$noc, region = noc_regions$region, flag = countrycode(noc_regions$region, 'country.name', 'unicode.symbol'))

olympics <- left_join(olympics, flags, by = "noc")

olympics_cty <- olympics %>%
  group_by(region, sport, date) %>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop")
```

```{r}
olympics_cty %>%
  filter(sport == "Gymnastics") %>%
  ggplot(aes(x = date, y = total_score, group = region)) +
  geom_line()
```

In our first visualization, we will plot `medals_total` on the y-axis and `date` on the x-axis with lines grouped by `country` and faceting by `sport`. We will select just a few sports of interest to compare and use selective color mapping for countries of interest and transparency to strongly highlight a few countries and make the remaining lines indistinguishable in the background. If this technique remains visually overwhelming after prototyping our plot, we will filter to select just a few countries of interest in each sport and color map only those.

#### Plot 2

```{r}
gdp <- read_csv("data/gdp-per-capita-maddison-2020.csv")

gdp %<>%
  janitor::clean_names()
```

```{r}
olympics_gdp <- olympics %>%
  group_by(region, date) %>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop") %>%
  rename(entity = region) %>%
  mutate(year = year(date)) %>%
  left_join(gdp, by = c("entity", "year"))
```

```{r}
olympics_gdp %>%
  filter(year >= 2010) %>%
  ggplot(aes(x = gdp_per_capita, total_score)) +
  geom_point() +
  facet_wrap(~date)
```

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
