---
title: Project title
author: by Mr. Palmer's Penguins
output: github_document
---

```{r load-packages, echo = FALSE}
# load packages here
suppressPackageStartupMessages(library(tidyverse))
library(showtext) # cool fonts


```

```{r load-data, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))

```

```{r set-options, echo = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.asp= 0.618, out.width = "80%")

```

## Introduction

(1-2 paragraphs)
Brief introduction to the dataset.
You may repeat some of the information about the dataset provided in the introduction to the dataset on the TidyTuesday repository, paraphrasing on your own terms.
Imagine that your project is a standalone document and the grader has no prior knowledge of the dataset.

## Does the optimal build of an Olympic athlete differ by sex or by sport? How have the trends in height and weight of Olympic athletes evolved over time?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

We are interested in exploring the physical characteristics of Olympic athletes. The Olympic Games is unique in showcasing a variety of disciplines and the body types of athletes who succeed at them. For example, imagine a typical male Olympic wrestler, versus a typical female gymnast. Pretty disparate builds, right? Even within one sport, say gymnastics, we hypothesize that there may be different optimal heights and weights for successful athletes of different genders  Therefore, to answer our first question, we will visualize the variables `height`, `weight`, `age`, `sex`, `sport`, and `year`.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.


For our first plot, we plan to use boxplots to plot the distribution of a size characteristic by `sex` (since characteristics tend to differ among males and females) and faceting by `sport`. We are unsure which of the `height`, `weight`, `height / weight`, or `bmi` ratio will be most telling when examined across sexes and sports. We will start by making 4 plots (each with one of these on the y-axis) and then select the plot where the trends are most insightful.

Two of these variables require new calculations in the data wrangling step. In order to plot the height-weight-ratio, we will have to create two new calculated columns. First, a `height_weight_ratio` column that simply divides an athlete's `height` by their `weight`, and a more complicated `BMI` column that represents an existing metric--[BMI](https://en.wikipedia.org/wiki/Body_mass_index)--that integrates height and weight into a single measure. . 

Based on the results from these preliminary plots, in addition to narrowing in on a size characteristic, we might pick a few select sports to focus in on (i.e. most popular sports) or group the sports based on common characteristics (ex: contact vs. non-contact, etc.) since there are 66 sports in the dataset.

### Analysis

``` {r wrangling-plot-one}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(height_weight_ratio = height / weight,
         BMI = 10000 * weight / (height * height))

# load Olympic font
font_add_google(name = "Oswald")
showtext_auto()
```

After examining each of `height`, `weight`, `height_weight_ratio`, and `BMI`, we have chosen to proceed with `BMI` as our size characteristic for this plot, as the visualizations were the most accessible and intuitive to interpret. After examining the plot displaying the `BMI` for all sports, we are choosing a few select groups of sports to include. The next step is to create separate dataframes for each of these categories. We also want to represent each athlete once on the plot, so we will have to use `distinct()`.

``` {r mutate-category}
# create variable to represent category of sport
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, .keep_all = TRUE) %>%
  mutate(sport = factor(sport, levels = rev(chosen_sports)),
         sex = factor(sex, labels = c("Female", "Male"))) %>%
  filter(sport %in% chosen_sports) %>%
  mutate(category = case_when(
    sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling") ~ "weightclass",
    sport %in% c("Gymnastics", "Trampolining") ~ "acrobatic",
    sport %in% c("Archery", "Shooting") ~ "coordination",
    TRUE ~ "diverse"
  ))

# create separate dataframes for each of these categories
olympics_weightclass <- olympics %>%
  filter(category == "weightclass")

olympics_acrobatic <- olympics %>%
  filter(category == "acrobatic")

olympics_coordination <- olympics %>%
  filter(category == "coordination")

olympics_diverse <- olympics %>%
  filter(category == "diverse")
```

We are now ready to create a separate plot for each sport category:

``` {r plot-weightclass}
ggplot(olympics_weightclass, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m^2)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "Of athletes competing in weightclass-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#E1AA7D", "#B6D094")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 9),
        legend.title = element_blank(),
        legend.position = c(0.82, 0.88),
        legend.direction = "horizontal",
        legend.background = element_rect(size = 0.3),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-acrobatic}
ggplot(olympics_acrobatic, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m^2)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "Of athletes competing in acrobatic Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#E1AA7D", "#B6D094")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 9),
        legend.title = element_blank(),
        legend.position = c(0.92, 0.98),
        #legend.direction = "horizontal",
        legend.background = element_rect(size = 0.3),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-coordination}
ggplot(olympics_coordination, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m^2)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "Of athletes competing in coordination-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#E1AA7D", "#B6D094")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 9),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        legend.background = element_rect(size = 0.3),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-diverse}
ggplot(olympics_diverse, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m^2)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "Of athletes competing in Olympic sports which encompass many skills",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#E1AA7D", "#B6D094")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 9),
        legend.title = element_blank(),
        legend.position = c(0.95, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.key.size = unit(0.5, "cm"))
```

We found a citation for the Tokyo 2020 Olympics logo font [here](https://www.reddit.com/r/identifythisfont/comments/4ig8ua/font_used_on_the_tokyo_2020_logo/) and were pointed to an open-source alternative [here](https://graphicdesign.stackexchange.com/questions/7178/is-there-a-din-font-free-alternative). We got the hex code used in plot text manually from that logo source. We chose the hex codes for our favorite gender color mapping from Telegraph 2018 [here](https://blog.datawrapper.de/gendercolor/).

We can now visualize these groups on the same plot, to get a better sense of broader variability in size characteristics among an array of Olympic sports.

``` {r fonts-and-plot-one, warning = FALSE}



ggplot(olympics, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m^2)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "Of athletes competing in selected Olympic sports from 1912-2020",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#E1AA7D", "#B6D094")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 9),
        legend.title = element_blank(),
        legend.position = c(0.82, 0.88),
        legend.direction = "horizontal",
        legend.background = element_rect(size = 0.3),
        legend.key.size = unit(0.5, "cm"))
  
```

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

In all plots, men typically have a higher-centered distribution of `BMI` than women in the same sport, but other trends are more apparent by `category`. In the first plot, which includes wrestling, weightlifting, judo, and boxing, we see a strong right skew.


gymnastics, trampolining (our original hypothesis; the boxes don't overlap! also the lower range of the female being like 12 makes me sad), 
shooting, archery (very wide range, makes sense as you body type doesn't really matter as much as a steady hand. shooting showing more disparity than archery, as well)
rowing, athletics (between coxswains in rowing, then sprinting vs throwing in althetics, no surprise that these are the top variables!)

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## Question 2 <- Update title to relate to the question you’re answering

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
