---
title: Project title
author: by Mr. Palmer's Penguins
output: github_document
---

```{r load-packages, echo = FALSE}
# load packages here
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(gganimate))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(gifski))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(showtext)) # cool fonts
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggtext)) # superscript in plot labels
```

```{r load-data, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
```

```{r set-options, echo = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.asp= 0.618, out.width = "80%")
```

## Introduction

The set, which includes `r ncol(olympics)` variables and `r nrow(olympics)` observations, provides insight into the athletes that competed in and the results of the Olympic games from Athens 1896 to Rio 2016. Each observation includes information about an athlete on a per event basis. Therefore, if an athlete competes in multiple different events, there will be multiple different observations reflecting each event an individual participated in. Each observation details medal results (Gold, Silver, Bronze, or NA), the athlete's background (age, sex, height, weight, team they're competing for, etc.), and context about the event (where and when it was held, season, sport the athlete competes in, etc.).

We chose this dataset because we think it can provide a lot of good insights into some of the trends at the Olympics over the past century as well as provide strong opportunities for our group to develop our visualization capabilities. 

This dataset was accessed on [kaggle.com](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results) and includes data scraped in May 2018 from Sports Reference/OlympStats, sports statistics sites, by Randi Griffin, a Senior Data Scientist at Boston Consulting Group. Moreover, it was selected as part of the [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021#readme7c70d95441aec295a1e92da9d71e2872877d663c) challenge on 7/27/21.

## Do certain physical characteristics (such as body mass index, age, and sex) of Olympic participants differ by sport and/or change with time?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

As we know, the Olympic Games is a unique opportunity for diverse and talented individuals to showcase their skills across a variety of disciplines. As a result, we are interested in better understanding how the diversity of physical characteristics in the set of competing athletes potentially differs by sport and/or has changed with time. Three physcial characteristics of interest are body mass index (a ratio of `height` and `weight`), `sex`, and `age`. It's often the case that the media/press discuss these these physical characteristics as key metrics when describing athletes. Therefore, we are curious about better understanding the extent to which these differ (if at all).

Firstly, we are hoping to understand if the optimal build of an Olympic athlete differs by `sex` or by `sport`. For example, a high performing male Olympic wrestler is likely to have  a different build versus a typical female gymnast. Pretty disparate builds, right? Even within one sport, say gymnastics, we hypothesize that there may be different optimal heights and weights for successful athletes of different sexes.

Secondly, we will assess if the mean age of Olympic participants for a given `year` has changed with time by `sex`. We know that over time the Olympics have transitioned from an amateur to professional competition. Therefore, we are curious to see if this has any association with the ages of participants since we know that certain age are more optimal for "peak performance" than others.
We will separate the visualization by `sex` because while there an approximately equal number of males and females in the world, we know that access to athletics participation is not always equal which could be associated with participation rates by `age`.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.


For our first plot, we plan to use boxplots to plot the distribution of a size characteristic by `sex` (since characteristics tend to differ among males and females) and faceting by `sport`. We are unsure which of the `height`, `weight`, `height / weight`, or `bmi` ratio will be most telling when examined across sexes and sports. We will start by making 4 plots (each with one of these on the y-axis) and then select the plot where the trends are most insightful.

Two of these variables require new calculations in the data wrangling step. In order to plot the height-weight-ratio, we will have to create two new calculated columns. First, a `height_weight_ratio` column that simply divides an athlete's `height` by their `weight`, and a more complicated `BMI` column that represents an existing metric--[BMI](https://en.wikipedia.org/wiki/Body_mass_index)--that integrates height and weight into a single measure. . 

Based on the results from these preliminary plots, in addition to narrowing in on a size characteristic, we might pick a few select sports to focus in on (i.e. most popular sports) or group the sports based on common characteristics (ex: contact vs. non-contact, etc.) since there are 66 sports in the dataset.

For our second plot, we are going to make a scatter plot of average `age` vs. `year` color mapped by `sex`. To start, we will find the average age of the participants by sex for each year. This will allow us to a not too crowded scatter plot since there are now only two observations (one for Male and one for Female) for each year. We choose a scatter plot because we believe it can best best represent the relationship and highlight potential trends. We considered a line plot but ultimately decided that the autonomy of each of the points (rather than them being connected) would better allow us to compare observations `year` by `year`. Since there is. unique context that surrounds each Olympics game, we feel as though a more year-specific approach could be more meaningful.

### Analysis

``` {r wrangling-plot-one}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(height_weight_ratio = height / weight,
         BMI = 10000 * weight / (height * height))

# load Olympic font
font_add_google(name = "Oswald")
showtext_auto()
```

After examining each of `height`, `weight`, `height_weight_ratio`, and `BMI`, we have chosen to proceed with `BMI` as our size characteristic for this plot, as the visualizations were the most accessible and intuitive to interpret. After examining the plot displaying the `BMI` for all sports, we are choosing a few select groups of sports to include. The next step is to create separate dataframes for each of these categories. We also want to represent each athlete once on the plot, so we will have to use `distinct()`.

``` {r mutate-category}
# create variable to represent category of sport
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  filter(sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling", "Gymnastics", "Trampolining", "Archery", "Shooting", "Athletics", "Rowing")) %>%
  mutate(category = case_when(
                    sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling") ~ "weightclass",
                    sport %in% c("Gymnastics", "Trampolining") ~ "acrobatic",
                    sport %in% c("Archery", "Shooting") ~ "coordination",
                    TRUE ~ "diverse"),
        category = factor(category, levels = c("acrobatic", "diverse", "coordination", "weightclass")))

# create separate dataframes for each of these categories
olympics_weightclass <- olympics %>%
  filter(category == "weightclass")

olympics_coordination <- olympics %>%
  filter(category == "coordination")

olympics_diverse <- olympics %>%
  filter(category == "diverse")

olympics_acrobatic <- olympics %>%
  filter(category == "acrobatic")
```

We are now ready to create a separate plot for each sport category. We will also silence messages which indicate that rows with NA have been dropped from the dataframe when creating the boxplots.

``` {r plot-weightclass, warning = FALSE}
ggplot(olympics_weightclass, mapping = aes(y = fct_rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in weightclass-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-coordination, warning = FALSE}
ggplot(olympics_coordination, mapping = aes(y = rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in coordination-based Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-diverse, warning = FALSE}
ggplot(olympics_diverse, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in Olympic sports which encompass many skills",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

``` {r plot-acrobatic, warning = FALSE}
ggplot(olympics_acrobatic, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in acrobatic Olympic sports",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

We found a citation for the Tokyo 2020 Olympics logo font [here](https://www.reddit.com/r/identifythisfont/comments/4ig8ua/font_used_on_the_tokyo_2020_logo/) and were pointed to an open-source alternative [here](https://graphicdesign.stackexchange.com/questions/7178/is-there-a-din-font-free-alternative). We got the hex code used in plot text manually from that logo source. We chose the hex codes for our favorite gender color mapping from Telegraph 2018 [here](https://blog.datawrapper.de/gendercolor/).

We can now visualize these groups on the same plot, to get a better sense of broader variability in size characteristics among an array of Olympic sports.

``` {r fonts-and-plot-one, warning = FALSE}
ggplot(olympics, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(x = "Body Mass Index (kg/m<sup>2</sup>)",
       y = NULL,
       color = "Sex",
       title = "Sex vs Body Mass Index Distribution",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",,
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.97),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
  
```

For our second plot, we will look at how `age` of Olympic participants by `sex` has changed with time.

We will re-load the CSV file to clear any manipulations done on the Olympics data frame to make the first plot.

```{r load-data-2, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
```

Since the observations in the data are done on a per event competed in basis (rather than a per athlete basis), we will use `distinct()` to filter the data so we only have unique athletes represented each `year`. In other words, each `year` will contain a set of observations representing one athlete each. This will remove cases where the age of an athlete who competed in multiple events for a given Olympics is repeated multiple times.

```{r distinct-year}
# keep only one observation per athlete
olympics <- olympics %>%
    distinct(id, year, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")))
```

Finally, we will make the plot showing the average yearly `age` of athletes by `sex`. We will drop observations in which the `age` is unknown.

```{r age-plot}
olympics %>%
  drop_na(age) %>%
  group_by(year, sex) %>%
  summarise(avg_age = mean(age)) %>%
  arrange(desc(sex)) %>%
  ggplot(mapping = aes(y = avg_age, x = year)) +
  geom_point(aes(color = sex)) +
  labs(x = "Year",
       y = "Age",
       color = "Sex",
       title = "Average Yearly Age of Athletes By Sex",
       subtitle = "For athletes competing in selected Olympic sports from 1912-2020",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB"), guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))
```

The mean was used, rather than the median, of `age` as this better captures the range of ages of athletes for a given year.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

In the "Sex vs Body Mass Index Distribution" plots, men typically have a higher-centered distribution of `BMI` than women in the same sport, but other trends are more apparent by `category`. In the first plot, which includes wrestling, weightlifting, judo, and boxing, we see a strong right skew in most distributions. We interpret this as a reflection of the weight classes present in those sports, with many participants entering in lower weight classes, giving a lower center, but a non-insignificant number of athletes with BMIs higher than 30. However, there are fewer at these high BMIs, possibly due to the difficulty in maintaining athletic competitiveness at that proportion. Shooting and archery also have generally higher BMIs than other sports, and again a right-skew is observed in the distributions. This reflects those sports' focus on hand-eye coordination, and thus athlete body type or height or weight are more variable.

The next group of sports, Athletics and Rowing, feature distributions with large variability on both sides. We interpret this as illustrating the diversity of body types considered desirable in this sports: from slender, lithe coxswains and sprinters to heavier, more powerful rowers and throwers. Finally, the largest disparities between `sex` in a particular sport are observed in Gymnastics and Trampoline. Here, we speculate that the female gymnastics disciplines favor petite, nimble athletes, as opposed to the more upper-body strength oriented male apparatuses, which contribute to a higher average BMI for male Olympic gymnasts.

Finally, for the "Average Yearly Age of Athletes By Sex" plot, it was found that the mean `age` of athletes by `sex` has changed pretty significantly over the years. For instance, the range of mean ages participating has converged pretty dramatically. Prior to the 1970s, the mean ages fluctuated in the 20-30 years old range with even some significant outliers hitting 35 and 50 years old in some cases. However, after the 1970s, it appears as though the mean ages converged in the 25-27 years old range. This makes sense because, according to a Britannica article found [here](https://www.britannica.com/sports/Olympic-Games), the 1970s were a turning point for the games moving from allowing professional athletes to compete rather than just amateurs. The article also notes that it wasn't until the 1980s that fully-fledged professional participation was permitted. With this information, we can speculate that there might be an association between the allowing of professional athletes and the convergence. Perhaps since most professional athletes are around 25 years old (a point of peak physical ability of many), the mean age reflects the demographic shift (amateur to professional) of participants. 

In addition to age converging, it is interesting to note the difference between Males and Females. For all years but 3, the mean age of males was higher than females. Perhaps since males take longer to physically mature (puberty is at a later age), that could explain why their ability to compete at "peak perform" in a professional context is at a later age. Moreover, it's possible that the extreme variability between mean ages of females prior to 1950 is the result of fewer females competing. With fewer competitors, it's more likely that outlier events (extremely old participants) can skew the mean. For example, prior to 1920, there were 210 female participants versus 9,790 male participants as depicted below.

```{r}
olympics %>%
  filter(year <= 1920) %>%
  count(sex)
```

These numbers are the total number of unique participants participating in each Olympics game prior to 1920.

## Question 2 <- Does the economic status of a country, measured by GDP, make it more successful than others in the Olympics? Which countries are most successful, regardless of GDP, and in which sports do they achieve this success?

### Introduction

The Olympics are a historic sports event that have been running for over a 125 years now, and to really grasp the full story of success at the Olympics, we want to first take a broader view and find out which countries have been most successful at all the Olympic games and in which sports. After conducting this research and creating relavant plots, we will continue on to testing a hypothesis we developed regarding a connection between a country's Olympic success and its economic output. We are interested in testing this theory that a more economically powerful country will have more success at the Olympics as we often hear that these countries are able to provide the resources and facilities needed for their athletes to achieve great success.

We will be using a variable we create, `medal_count`, for the first plot, and this will help us differentiate total Olympic success for different countries. We will then zoom in and look at one specific country's success, using the `sport`, `noc`, and `year` variables along with `medal_count`. For the second plot, we will load in external data regarding the GDP of countries that have participated in the Olympics, and use this external data along with `medal_count` and `medal_score` to compare success to GDP. `Medal_score` will be another variable we create that weights the winning of a gold medal heavier than a silver medal, and a silver heavier than a bronze.

### Approach

In our first visualization, we will create a bar chart to represent the winningest countries in the history of the Olympics. Country codes will be on the x-axis and total medal count, a new variable we will create, will be on the y-axis. To avoid having an extremely left skewed bar chart with many countries who have won very few medals, we will represent the top 10 winningest countries of all time on our bar chart. This plot will help us gauge the amount of medals the top ten most winningest countries have won, and how they compare to each other. After creating this plot, we will select one country of interest and examine which sports have contributed most heavily to this country's success. We will then plot out the results of these sports over time to see how their contribution to this country's medal count may have changed over the years.

For our second visualization,

### Analysis

Above we have loaded the olympic dataset and also the full country names into the noc_regions object.

```{r wrangling-medals}
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
noc_regions <- read_csv(file = paste0(here::here(), "/data/noc_regions.csv"))

# make appropriate date
olympics %<>%
  mutate(date = ifelse(season == "Winter", paste0(year, "-02-01"), paste0(year, "-07-01")),
         date = ymd(date),
         medal_winner = ifelse(is.na(medal), 0, 1),
         medal_score = case_when(
           medal == "Bronze" ~ 1,
           medal == "Silver" ~ 2,
           medal == "Gold" ~ 3,
           TRUE ~ 0))

noc_regions %<>%
  janitor::clean_names() %>%
  mutate(region = ifelse(region == "Boliva", "Bolivia", region))
```

The data wrangling step above involves adding the month and day to the dates of our summer and winter olympic data to help us plot it as time series data, should we need to do this. We then clean the noc_regions data.


```{r}
olympics_filter <- olympics %>%
  filter(season == "Summer") %>%
  group_by(noc) %>%
  dplyr::summarise(total_winners = sum(medal_winner), total_score = sum(medal_score))%>%
  filter(total_winners >= 910)

```

We have now filtered the data to be used in our first visualization for only Summer Olympic success, as different countries may be successful in different seasons. For this visualization, we are interested in the Summer Olympics. We then summarize the total medals as the sum of the `medal_winner` column, which represents 1 or 0 for if an athlete won a medal or not, and add up the `medal_score` column to get a total score that weights gold, silver and bronze medals.

###Plot 1

```{r}

ggplot(olympics_filter, aes(x = reorder(noc, total_winners), y = total_winners, fill = noc, color = noc))+
  geom_bar(stat='identity')+
  labs(x = "Country",
       y = "Medal Count",
       title = "All-time Summer Olympic Medal Counts",
       subtitle = "For the top 10 most winningest countries in history",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  scale_color_manual(values = c("#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB")) +
  scale_fill_manual(values = c("#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB","#5CC0AB")) +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.position = "none")


```

Here is our first plot, which represents the 10 countries that have won the most medals in Summer Olympic history.

```{r}

usa_fil <- olympics %>%
  filter(noc == "USA") %>%
  group_by(sport, year)%>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score))%>%
  filter(sport %in% c("Swimming", "Athletics"))
```

After taking a closer look at the data used to create the first visualization, we decided to represent the contribution of the USA's two most successful sports to our medal count at each Summer olympic games over time. We filter by the country code "USA" and by sport and year, filtering for only swimming and athletics.

```{r}

ggplot(usa_fil, aes(x = year, y = total_winners, color = sport, group = sport))+
  geom_line()+
  geom_point()+
  labs(x = "Year",
       y = "Medal Count",
       title = "USA Olympic Medals Won in Athletics and Swimming",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  scale_color_manual(values = c("#7B38EC","#5CC0AB"))+
  scale_x_continuous(limits = c(1896,2016), breaks = seq(1896, 2016, by = 8))+
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.18),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"))

```

Here we have plotted the number of medals won in Athletics and Swimming at each Summer Olympic games over the years.

#### Plot 2

```{r}
detach(package:plyr, unload = TRUE)
gdp <- read_csv("data/gdp-per-capita-maddison-2020.csv")

gdp %<>%
  janitor::clean_names() %>%
  mutate(entity = case_when(
    entity == "United States" ~ "USA",
    entity == "United Kingdom" ~ "UK",
    entity == "Congo" ~ "Republic of Congo",
    entity == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo",
    entity == "Cote d'Ivoire" ~ "Ivory Coast",
    entity == "Trinidad and Tobago" ~ "Trinidad",
    entity == "North Macedonia" ~ "Macedonia",
    entity == "Czechia" ~ "Czech Republic",
    TRUE ~ entity
  ))
```

```{r}
olympics_gdp <- olympics %>%
  group_by(region, date, season, flag) %>%
  summarize(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop") %>%
  rename(entity = region) %>%
  mutate(year = year(date),
         summer_highlight = ifelse(total_winners > 100, T, F),
         winter_highlight = ifelse(total_winners > 30, T, F)) %>%
  left_join(gdp, by = c("entity", "year")) %>%
  rename(country = entity)
```

```{r fig.asp = 1.2}
winter_highlight <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(2000:2020), total_winners > 30)

olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(2000:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = winter_highlight, size = winter_highlight), alpha = .5, size = .75, show.legend = FALSE) +
  geom_text_repel(data = winter_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
                                        color = "#7B38EC", size = 3.5, family = "Oswald") +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(title = "Total Medals Won vs. GDP per capita",
       subtitle = "Winter Olympics, 2002-2014",
       y = "Total Medals Won",
       x = "GDP per capita (USD)",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14))
```

```{r fig.asp = 1.2}
summer_highlight <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Summer", year %in% c(2004:2020), total_winners > 100)

olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Summer", year %in% c(2004:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = summer_highlight, size = summer_highlight), 
             alpha = .5, size = .75, show.legend = FALSE) +
  geom_text_repel(data = summer_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
                                        color = "#7B38EC", size = 3.5, family = "Oswald") +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(title = "Total Medals Won vs. GDP per capita",
       subtitle = "Winter Olympics, 2004-2016",
       y = "Total Medals Won",
       x = "GDP per capita (USD)",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_y_continuous(breaks = seq(from = 0, to = 300, by = 100)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14))

olympics_gdp %>%
  filter(!is.na(gdp_per_capita), country != "Qatar", season == "Summer", year %in% c(2004:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = summer_highlight, size = summer_highlight), 
             alpha = .5, size = .75, show.legend = FALSE) +
  geom_text_repel(data = summer_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
                                        color = "#7B38EC", size = 3.5, family = "Oswald") +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(title = "Total Medals Won vs. GDP per capita",
       subtitle = "Winter Olympics, 2004-2016",
       y = "Total Medals Won",
       x = "GDP per capita (USD)",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_y_continuous(breaks = seq(from = 0, to = 300, by = 100)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14))
```



```{r cache = TRUE}
winter_highlight <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(1950:2020), total_winners > 30)

winter_gdp <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(1950:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = winter_highlight), alpha = .5, size = .75, show.legend = FALSE) +
  geom_text(data = winter_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
                                        color = "#7B38EC", size = 4, family = "Oswald", vjust = .2) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(title = "Total Medals Won vs. GDP per capita",
       subtitle = "Winter Olympics, 1950-2020\nYear: {closest_state}",
       y = "Total Medals Won",
       x = "GDP per capita (USD)",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")) +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.68),
        legend.background = element_rect(size = 0.3),
        legend.margin = margin(1, 5, 5, 5),
        legend.key.size = unit(0.5, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260", size = 10),
        axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
        legend.text = element_text(family = "Oswald", color = "#092260", size = 14)) +
  transition_states(year, transition_length = 1, state_length = .3)

animate(winter_gdp, duration = 20, fps = 10, renderer = gifski_renderer())
```

### Discussion

After plotting the 10 most successful Summer Olympic countries in history, it is clear that the USA has a wide lead over the USSR, which is in second place, and other countries behind them, including Great Britain, Germany, France, and Italy. The United States, as seen in the visualization, has won over 5,000 Summer Olympic medals, while the USSR has won just over 2,000, Britain a similar amount, and we can see a steady decline after that. We can assume that the USSR's grasp on second place total medals will soon be overtaken as they have split into many smaller countries and Russia, and can no longer win medals under the USSR title. Because of this and the wide gap between first and second place, we decided to examine the data to see which sports contributed most to the success of the Americans. After examining the data and grouping the USA's medals by sport, it was clear that the two most successful American Summer Olympic sports were swimming and athletics.

The immense contribution of swimming and athletics to American Summer Olympic success makes sense as these two sports consist of many events with lots of medals up for grabs, as compared to a sport like basketball in which only one medal can be one at each Summer Olympics. We wanted to now examine the success of the USA in these two sports, swimming and athletics, and see if it has changed over time. After grouping USA's Summer Olympic data by sport and year, we filter for swimming and athletic data to see how the Americans have fared in these two sports over time, compared to each other. The plot shows that from the first Olympic games until the 1960 games, the USA won more athletic medals than swimming medals at every occurrence of the Summer Olympics. But, starting in 1960, The number of medals won by the USA in swimming overtook the same number for athletics, and since then, the United States has won more swimming than athletic medals at every Olympic games. The data is missing values for 1940 and 1944 because of World War II, and for 1980 because of the worldwide boycott of the Russian Olympic games. The trend seen is very interesting, as we can see that in the late 20th century, there was a small gap between the number of medals won in swimming and athletics, but starting in 1996 and continuing through the turn of the century, the gap started to widen. This could be because of a multitude of reasons, including more USA swimming development or the addition of more swimming events. In the last three Olympics for which we have data, we can see that the gap between the two sports stays relatively steady but success in both is on the rise.

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
