---
title: "Project proposal"
author: "Mr. Palmer's Penguins"
output: github_document
---
 
```{r load-packages-and-data, message = FALSE, warning = FALSE}
library(tidyverse)

# load data from tidytuesday site and write to csv in data folder
readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv") %>%
  write_csv(file = paste0(here::here(), "/data/olympics_data.csv"))

# read data from csv
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
```

## Dataset

This dataset was accessed on Kaggle.com and includes data scraped in May 2018 from Sports Reference/OlympStats, sports statistics sites, by Randi Griffin, a Senior Data Scientist at Boston Consulting Group. Moreover, it was selected as part of the TidyTuesday challenge on 7/27/21. Citation information is included at the bottom of this section.

The set, which includes `r ncol(olympics)` variables and `r nrow(olympics)` observations, provides insight into the athletes that competed in and the results of the Olympic games from Athens 1896 to Rio 2016. Each observation includes information about an athlete on a per event basis. Therefore, if an athlete competes in multiple different events, there will be multiple different observations reflecting each event an individual participated in. Each observation details medal results (Gold, Silver, Bronze, or NA), the athlete's background (age, sex, height, weight, team, etc.), and context about the event (where and when it was held, season, etc.).

The full list of variables are included here and described below:

```{r load-csv, message = FALSE, warning = FALSE}
glimpse(olympics)
```

`ID` - This is an athlete-specific number. Each athlete is assigned one unique identifying number in the dataset

`Name` - The name of the athlete competing.

`Sex` - Indicated whether the athlete is Male (M) or Female (F)

`Age` - The age of the athlete in years

`Height` - The height of the athlete in centimeters

`Weight` - The weight of the athlete in kilograms

`Team` - The team (same as the country in most cases) the athlete represents

`NOC` - The 3-letter National Olympic Committee code associated with each team (indicative of country in most cases)

`Games` - The year (1896-2016) and season (Summer or Winter) in which the event took place. This is a character combination of `Year` and `Season`. It is important to note that the Summer and Winter games currently alternate every two years, but used to take place in the same year.

`Year` - The year (1896-2016) in which the event took place. 

`Season` - The season (Summer or Winter) in which the event took place.

`City` - The city in which the Olympic games for that year/season were hosted.

`Sport` - The sport the athlete is competing in.

`Event` - The event the athlete is competing in within a sport.

`Medal` - Indicates whether a Gold, Silver, or Bronze medal was won. Responses can also take the form NA if no medal was won or there is incomplete information.

We chose this dataset because we think it can provide a lot of good insights into some of the trends at the Olympics over the past century as well as provide strong opportunities for our group to develop our visualization capabilities. 

These data offer many avenues to explore when analyzing, such as medalwinning success on an athlete-level scale, on a country-level scale, or for a particular sport over time, etc. Complementing this variety of potential analysis paradigms is the medley of variable types. Several variable types such as categorical, ordinal, time-series, and ratio are well represented in the data. This also affords us the opportunity to calculate new variables, such as converting the medal variable into a logical medal winner or non-medal winner.

Finally, aside from the properties of the data itself, we chose to analyze data from the Olympic games due to the recency of the 2020/2021 Tokyo Games and the role of the Games in reflecting the global culture of a time period. Historic events like the 1980 Moscow Games that were boycotted by the US and many other countries, the several instances of Games that were canceled during the World Wars, and the late 1980s advent of professional superstar Olympians all potentially provide insights into the story told by the data.

#### Sources

[TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021#readme7c70d95441aec295a1e92da9d71e2872877d663c)

[Kaggle](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results)

## Questions

The two questions you want to answer.

1. How have the physical characteristics (height-to-weight ratio, age, and sex) of Olympic participants changed over time and do they differ by sport?

Variables: `height`, `weight`, `age`, `sex`, `sport`, `year`

2. How has the success of different countries in a particular sport changed with time? 

Variables: `year`, `NOC`, `medal count`, `sport`, `country`

## Analysis plan

Here is a plan for answering each of the questions:

### For Question 1

We are interested in exploring the physical characteristics of Olympic athletes. Therefore, to answer our first question, we will visualize the variables `height`, `weight`, `age`, `sex`, `sport`, and `year`. 

#### Plot 1:

We are all used to hearing that Olympic athletes have certain favorable physical condition, but those characteristics differ by sport. For example, volleyball and basketball players are usually exceptionally tall, while wrestlers are usually shorter and built. Therefore, for our first plot, we plan to explore how size characteristics (either using `height`, `weight`, or the `height:weight` ratio) differ by `sport` and `sex`. In other words, we plan to use boxplots to plot the distribution of a size characteristic by `sex` (since characteristics tend to differ by division) and faceting by `sport`.

Since we are unsure which of the `height`, `weight`, or the `height:weight` ratio will be most telling, we will start by making 3 plots (each with one of these on the y-axis) and then select the plot where the trends are most insightful.

In order to plot the height-weight-ratio, we will have to create a new variable and divide height by weight. 

Based on the results, in addition to narrowing in on a size characteristic, we might pick a few select sports to focus in on (i.e. most popular sports) or group the sports based on common characteristics (ex: contact vs. non-contact, etc.) since there are 66 sports in the dataset.

#### Plot 2:

For our second plot, we will look more closely at how size characteristics have changed over time, using `year` to make a time series line plots for these different values. In other words, we will plot one of the size characteristics (`height`, `weight`, or the height:weight ratio) on the y-axis vs. time on the x-axis. Lines by `sex` will connect the observations. We might also facet these plots by `age` to see if, for example, height of athletes has changed more or less over time for younger athletes vs. older athletes. In order to facet by `age`, we will have to make `age` a categorical variable by setting different cutoffs in which we group observations. 

The ages in the dataset range from 10 to 97 years old. Therefore, we will plot a histogram and do summary statistics to understand which cutoffs make the most sense given the data. Since athletes tend to be in their mid 20s, we'll likely have smaller differences in ages in the 20-30 age range whereas all athletes 40-97, for example, might be grouped together. However, as mentioned, this would depend on the distribution of the data.

Like in Plot 1, we will evaluate which of `height`, `weight`, or the `height:weight` ratio provides the most insightful visualization. Ideally, the variable we choose in Plot 1 will be the same as in Plot 2; however, we understand creating visualizations is an iterative process are open to evaluating all options as we plot.

### For Question 2

#### Data wrangling

The original variables we will use to answer this question are `noc`, `medal`, `sport` (perhaps `event` if some higher granularity proves informative rather than overwhelming), and `year` and `season`. Both `year` and `season` are important for placing each Olympics in time and will require some cleaning in conversion to an accurate variable `date`. Many of us are familiar with even recent anomalies in Olympic labeling (such as these recent 2020 Summer Olympics, which took place in 2021). The exact begin and end dates of each Olympics are not the goal of this step. We simply wish to be able to correctly order the games chronologically, which may require more information than just the year and season in some cases.

Next, we will mutate `medal` from a categorical variable with 'Gold', 'Silver', 'Bronze', and NA as outcomes to either a numeric indicator (medal = 1 for any win and medal = 0 for no medal received) or a score that rewards the most points for Gold, then Silver, then Bronze. Through iterative visualization we will determine whether examining success only based on medal count or differentiating gold, silver, and bronze victories allows for a more meaningful display of trends. We will also later group the observations by region of National Olympic Committee, or `noc`, which reflects what we understand to be country identities today more accurately than `team` does. Merging in this [secondary dataset](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results?select=noc_regions.csv) provided with the athlete-event level data by Kaggle will allow us to expand our country labeling from three-letter NOC abbreviations to complete country names. Additionally, the dataframe in the package [countrycode](https://vincentarelbundock.github.io/countrycode/) links countries by their names in various formats to their unicode flag emojis. Merging this data in will assist us with our visualizations of this question.

After the above merging and necessary cleaning, we will group the data by `country`, `sport`, and `date` to compute a new variable called `medals_total` by taking the sum of either our medal-won indicator or our medal score assigned to differentiate gold, silver and bronze victories. These steps should allow us to begin an initial approach to visualization.

#### Plot 1:

In our first visualization, we will plot `medals_total` on the y-axis and `date` on the x-axis with lines grouped by `country` and faceting by `sport`. We will select just a few sports of interest to compare and use selective color mapping for countries of interest and transparency to strongly highlight a few countries and make the remaining lines indistinguishable in the background. If this technique remains visually overwhelming after prototyping our plot, we will filter to select just a few countries of interest in each sport and color map only those.

#### Plot 2:

Our second visualization will require historical GDP per capita data by country. We are interested in attempting to understand how the nationwide economic factors may be associated with trends in countries' Olympic performance over time. The external data we have chosen to merge in comes from the Maddison Project Database, a compilation of researchers' work estimating economic growth for individual countries ([Bolt and van Zanden, 2020](https://ourworldindata.org/grapher/gdp-per-capita-maddison-2020)).

We will create a scatterplot with `gdp_per_capita` on the x axis and `medals_total` on the y axis. We plan to highlight interesting outliers with country emojis, and we will facet by `season` only if that reveals an interesting difference between the economic indicator-athletic performance association. We will also use an iterative approach to determine whether selecting one or a few year(s) of interest to maintain a one-to-one ratio of points to countries or plotting all the data points (one for each country-year combination) will appear more informative to the viewer. Alternatively, if we are permitted to produce an animated plot and animation is feasible given our other aesthetic goals, we will animate the plot with each frame representing a single Olympics and year (or Olympics and season if faceting by season proves to be uninformative).

#### A note on faceting

It is important to note that for several of our proposed plots, we mention mapping categorical variables with many levels--such as `country` or `sport`--to visualization tools like faceting and `geom_line()`. It is crucial to clarify that we will carefully, intentionally choose a subset of possible countries or sports to visualize in each of these cases (based on population size, medal count, compelling trends, etc.). We acknowledge that failing to do so may lead to messy plots that are difficult to interpret, and thus will prioritize legibility and meaning in our ultimate visualizations.
