---
title: "Project proposal"
author: "Mr. Palmer's Penguins"
output: github_document
---

```{r load-packages-and-data, message = FALSE, warning = FALSE}
library(tidyverse)

# load data from tidytuesday site and write to csv in data folder
readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv") %>%
  write_csv(file = paste0(here::here(), "/data/olympics_data.csv"))

# read data from csv
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"))
```

## Dataset

This dataset was accessed on Kaggle.com and includes data scraped in May 2018 from Sports Reference/OlympStats, sports statistics sites, by Randi Griffin, a Senior Data Scientist at Boston Consulting Group. Moreover, it was selected as part of the TidyTuesday challenge on 7/27/21. Citation information is included at the bottom of this section.

The set, which includes `r ncol(olympics)` variables and `r nrow(olympics)` observations, provides insight into the athletes that competed in and the results of the Olympic games from Athens 1896 to Rio 2016. Each observation includes information about an athlete on a per event basis. Therefore, if an athlete competes in multiple different events, there will be multiple different observations reflecting each event an individual participated in. Each observation details medal results (Gold, Silver, Bronze, or NA), the athlete's background (age, sex, height, weight, team, etc.), and context about the event (where and when it was held, season, etc.).

The full list of variables are included here and described below:

```{r load-csv, message = FALSE, warning = FALSE}
glimpse(olympics)
```

`ID` - This is an athlete-specific number. Each athlete is assigned one unique identifying number in the dataset

`Name` - The name of the athlete competing.

`Sex` - Indicated whether the athlete is Male (M) or Female (F)

`Age` - The age of the athlete in years

`Height` - The height of the athlete in centimeters

`Weight` - The weight of the athlete in kilograms

`Team` - The team (same as the country in most cases) the athlete represents

`NOC` - The 3-letter National Olympic Committee code associated with each team (indicative of country in most cases)

`Games` - The year (1896-2016) and season (Summer or Winter) in which the event took place. This is a character combination of `Year` and `Season`. It is important to note that the Summer and Winter games currently alternate every two years, but used to take place in the same year.

`Year` - The year (1896-2016) in which the event took place. 

`Season` - The season (Summer or Winter) in which the event took place.

`City` - The city in which the Olympic games for that year/season were hosted.

`Sport` - The sport the athlete is competing in.

`Event` - The event the athlete is competing in within a sport.

`Medal` - Indicates whether a Gold, Silver, or Bronze medal was won. Responses can also take the form NA if no medal was won or there is incomplete information.

We chose this dataset because we think it can provide a lot of good insights into some of the trends at the Olympics over the past century as well as provide strong opportunities for our group to develop our visualization capabilities. 

These data offer many avenues to explore when analyzing, such as medalwinning success on an athlete-level scale, on a country-level scale, or for a particular sport over time, etc. Complementing this variety of potential analysis paradigms is the medley of variable types. Several variable types such as categorical, ordinal, time-series, and ratio are well represented in the data. This also affords us the opportunity to calculate new variables, such as converting the medal variable into a logical medal winner or non-medal winner.

Finally, aside from the properties of the data itself, we chose to analyze data from the Olympic games due to the recency of the 2020/2021 Tokyo Games and the role of the Games in reflecting the global culture of a time period. Historic events like the 1980 Moscow Games that were boycotted by the US and many other countries, the several instances of Games that were canceled during the World Wars, and the late 1980s advent of professional superstar Olympians all potentially provide insights into the story told by the data.

Source: TidyTuesday

https://github.com/rfordatascience/tidytuesday/tree/master/data/2021#readme7c70d95441aec295a1e92da9d71e2872877d663c

Source: Kaggle

https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results 

## Questions

The two questions you want to answer.

1. How have the physical characteristics (height-to-weight ratio, age, and sex) of Olympic participants changed over time and do they differ by sport?

Variables: `height`, `weight`, `age`, `sex`, `sport`, `year`

2. How has the success of different countries in a particular sport changed with time? 

Variables: `year`, `NOC`, `medal count`, `sport`

## Analysis plan

Here is a plan for answering each of the questions:

### For Question 1

To answer our first question, we will evaluate the variables `height`, `weight`, `age`, `sex`, `sport`, and `year`. After looking at the dataset, we are interested in exploring the physical characteristics of Olympic athletes. We are all used to hearing that these athletes are in peak physical condition, but this classification differs by sport and has probably changed over the years.

We will start by creating a new variable, `height-weight-ratio`. We are interested in exploring the height to weight ratio for different sports in the Olympics, as different sports require different body-types for success. For example, volleyball and basketball players are usually exceptionally tall, while wrestlers are usually shorter and built. We will create this variable and use it in our exploration of the data to see if we can find anything interesting.

We now want to see how this ratio and other characteristics, like height and weight individually, differ in different sports. We can accomplish this by plotting `height` and `weight`, maybe with boxplots, and faceting by `sport`. Based on the results, we might pick a few select sports to focus in on or group the sports based on common characteristics since there are many (probably too many for one plot) to facet by. Moreover, we also plan to analyze this information by `sex` since characteristics can differ for each division.

To see how these values have changed over the years, we will then use `year` to make time series line plots for these different values. We can also facet these plots by `sport` to see if, for example, height of athletes has changed more or less over time in one sport as compared to another. Another part of the question we may be able to answer pertains to the effect of athletes' age on physical characteristics. We can plot `height` and `weight` against `age` as a continuous variable, or we can transform `age` into a discrete, factor variable and create boxplots for different characteristics at different age levels.

### For Question 2

The original variables we will use to answer this question are `noc`, `medal`, `sport` (perhaps `event` if some higher granularity proves informative rather than overwhelming), and `year` and `season`. Both `year` and `season` are important for placing each Olympics in time and will require some cleaning in conversion to an accurate variable `date`. Many of us are familiar with even recent anomalies in Olympic labeling (such as these recent 2020 Summer Olympics, which took place in 2021).

Next, we will mutate `medal` from a categorical variable with 'Gold', 'Silver', 'Bronze', and NA as outcomes to either a numeric indicator (medal = 1 for any win and medal = 0 for no medal received) or a score that rewards the most points for Gold, then Silver, then Bronze. Through iterative visualization we will determine whether examining success only based on medal count or differentiating gold, silver, and bronze victories allows for a more meaningful display of trends. We will also later group the observations by region of National Olympic Committee, or `noc`, which reflects what we understand to be country identities today more accurately than `team` does. Merging in this [secondary dataset](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results?select=noc_regions.csv) provided with the athlete-event level data by Kaggle will allow us to expand our country labeling from three-letter NOC abbreviations to complete country names. Additionally, the dataframe in the package [countrycode](https://vincentarelbundock.github.io/countrycode/) links countries by their names in various formats to their unicode flag emojis. Merging this data in will assist us with our visualizations of this question.

After the above merging and necessary cleaning, we will group the data by `country`, `sport`, and `date` to compute a new variable called `medals_total`. These steps should allow us to begin an initial approach to visualization, for example, by plotting `medals_total` on the y-axis and `date` on the x-axis with lines grouped by `country` and faceting by `sport`.

It is important to note that for several of our proposed plots, we mention mapping categorical variables with many levels--such as `country` or `sport`--to visualization tools like facetting and `geom_line()`. It is crucial to clarify that we will carefully, intentionally choose a subset of possible countries or sports to visualize in each of these cases (based on population size, medal count, compelling trends, etc.). We acknowledge that failing to do so may lead to messy plots that are difficult to interpret, and thus will prioritize legibility and meaning in our ultimate visualizations.
