---
title: "Beyond the Rings:"
subtitle: "Analyzing Olympic Athlete and Medal-Winning Country Characteristics"
author: "Mr. Palmer's Penguins <br> Sarab Bhasin, Lillian Clark, Evan Dragich, Drew Flanagan"
institute: "Duke University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
# Remove any packages from this list that you're not using
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(gganimate))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(gifski))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(showtext)) # cool fonts
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggtext)) # superscript in plot labels
suppressPackageStartupMessages(library(styler))
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r load-data, include=FALSE}
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
```

```{r style-slides, echo = FALSE}
style_xaringan(
  title_slide_background_image = "https://upload.wikimedia.org/wikipedia/commons/thumb/5/5c/Olympic_rings_without_rims.svg/1200px-Olympic_rings_without_rims.svg.png"
)
```

class: inverse, middle, center
# Do certain physical characteristics (such as body mass index, age, and sex) of Olympic participants differ by sport and/or change with time?

---
class: middle, center
``` {r wrangling-plot-one, echo = FALSE}
# Create new calculated variables to represent physical build
olympics <- olympics %>%
  mutate(
    height_weight_ratio = height / weight,
    BMI = 10000 * weight / (height * height)
  )

# load Olympic font
font_add_google(name = "Oswald")
showtext_auto()
```

``` {r mutate-category, echo = FALSE}
# create variable to represent category of sport
# keep only one observation per athlete
olympics <- olympics %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  filter(sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling", "Gymnastics", "Trampolining", "Archery", "Shooting", "Athletics", "Rowing")) %>%
  mutate(
    category = case_when(
      sport %in% c("Boxing", "Judo", "Weightlifting", "Wrestling") ~ "weightclass",
      sport %in% c("Gymnastics", "Trampolining") ~ "acrobatic",
      sport %in% c("Archery", "Shooting") ~ "coordination",
      TRUE ~ "diverse"
    ),
    category = factor(category, levels = c("acrobatic", "diverse", "coordination", "weightclass"))
  )

# create separate dataframes for each of these categories
olympics_weightclass <- olympics %>%
  filter(category == "weightclass")

olympics_coordination <- olympics %>%
  filter(category == "coordination")

olympics_diverse <- olympics %>%
  filter(category == "diverse")

olympics_acrobatic <- olympics %>%
  filter(category == "acrobatic")
```

``` {r plot-weightclass, warning = FALSE, echo = FALSE, out.width = "90%"}
ggplot(olympics_weightclass, mapping = aes(y = fct_rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(
    x = "Body Mass Index (kg/m<sup>2</sup>)",
    y = NULL,
    color = "Sex",
    title = "Sex vs Body Mass Index Distribution",
    subtitle = "For athletes competing in weightclass-based Olympic sports", ,
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.9, 0.67),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```
---
class: middle, center
``` {r plot-coordination, warning = FALSE, echo = FALSE, out.width = "90%"}
ggplot(olympics_coordination, mapping = aes(y = rev(sport), x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(
    x = "Body Mass Index (kg/m<sup>2</sup>)",
    y = NULL,
    color = "Sex",
    title = "Sex vs Body Mass Index Distribution",
    subtitle = "For athletes competing in coordination-based Olympic sports", ,
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.9, 0.25),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```
---
class: middle, center
``` {r plot-diverse, warning = FALSE, echo = FALSE, out.width = "90%"}
ggplot(olympics_diverse, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(
    x = "Body Mass Index (kg/m<sup>2</sup>)",
    y = NULL,
    color = "Sex",
    title = "Sex vs Body Mass Index Distribution",
    subtitle = "For athletes competing in Olympic sports which encompass many skills", ,
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.9, 0.9),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```
---
class: middle, center
``` {r plot-acrobatic, warning = FALSE, echo = FALSE, out.width = "90%"}
ggplot(olympics_acrobatic, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(
    x = "Body Mass Index (kg/m<sup>2</sup>)",
    y = NULL,
    color = "Sex",
    title = "Sex vs Body Mass Index Distribution",
    subtitle = "For athletes competing in acrobatic Olympic sports", ,
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.9, 0.77),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```
---
class: middle, center
``` {r fonts-and-plot-one, warning = FALSE, echo = FALSE, out.width = "90%"}
ggplot(olympics, mapping = aes(y = sport, x = BMI, color = sex)) +
  geom_boxplot(position = position_dodge2(10)) +
  labs(
    x = "Body Mass Index (kg/m<sup>2</sup>)",
    y = NULL,
    color = "Sex",
    title = "Sex vs Body Mass Index Distribution",
    subtitle = "For athletes competing in selected Olympic sports from 1912-2020", ,
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  facet_grid(category ~ ., scales = "free_y", space = "free") +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260"),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.9, 0.77),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```

---
class: middle, center
```{r load-data-2, echo = FALSE}
# load data here
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
```

```{r distinct-year, echo = FALSE}
# keep only one observation per athlete
olympics <- olympics %>%
  distinct(id, year, .keep_all = TRUE) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male")))
```

```{r age-plot, echo = FALSE, out.width = "90%"}
olympics %>%
  drop_na(age) %>%
  group_by(year, sex) %>%
  dplyr::summarise(avg_age = mean(age), .groups = "drop_last") %>%
  arrange(desc(sex)) %>%
  ggplot(mapping = aes(y = avg_age, x = year)) +
  geom_point(aes(color = sex)) +
  labs(
    x = "Year",
    y = "Age",
    color = "Sex",
    title = "Average Yearly Age of Athletes By Sex",
    subtitle = "For athletes competing in selected Olympic sports from 1912-2020",
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB"), guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_text(family = "Oswald", color = "#092260", size = 14),
    axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260"),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.8, 0.68),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```

---
class: inverse, middle, center
# Does the economic status of a country, measured by GDP, make it more successful than others in the Olympics? Which countries are most successful, regardless of GDP, and in which sports do they achieve this success?
---
class: middle, center

```{r wrangling-medals, echo = FALSE}
# reload the Olympic dataset and also the full country names into the noc_regions object
olympics <- read_csv(file = paste0(here::here(), "/data/olympics_data.csv"), show_col_types = FALSE)
noc_regions <- read_csv(file = paste0(here::here(), "/data/noc_regions.csv"), show_col_types = FALSE)

# make appropriate date for chronological plotting and convert characters in medal wins to numeric values
olympics %<>%
  mutate(
    date = ifelse(season == "Winter", paste0(year, "-02-01"), paste0(year, "-07-01")),
    date = ymd(date),
    medal_winner = ifelse(is.na(medal), 0, 1),
    medal_score = case_when(
      medal == "Bronze" ~ 1,
      medal == "Silver" ~ 2,
      medal == "Gold" ~ 3,
      TRUE ~ 0
    )
  )

noc_regions %<>%
  janitor::clean_names() %>%
  # fix typo
  mutate(region = ifelse(region == "Boliva", "Bolivia", region))

olympics <- left_join(olympics, noc_regions, by = "noc")

olympics_filter <- olympics %>%
  filter(season == "Summer") %>%
  group_by(region) %>%
  dplyr::summarise(total_winners = sum(medal_winner), total_score = sum(medal_score)) %>%
  filter(total_winners >= 910)
```

```{r cumulative-medal-count-plot, echo = FALSE, out.width = "90%"}
ggplot(olympics_filter, aes(x = reorder(region, total_winners), y = total_winners)) +
  geom_bar(stat = "identity", fill = "#5CC0AB") +
  labs(x = "Country",
       y = "Medal Count",
       title = "All-time Summer Olympic Medal Counts",
       subtitle = "For the top 10 most winningest countries in history",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.spacing.y = unit(0.4, "cm"),
        plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
        plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
        plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
        axis.text.x = element_text(family = "Oswald", color = "#092260"),
        axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
        axis.text.y = element_text(family = "Oswald", color = "#092260"),
        axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
        legend.position = "none")
```
---
class: middle, center

```{r usa-victories-plot, echo = FALSE, out.width = "90%"}
usa_fil <- olympics %>%
  filter(noc == "USA") %>%
  group_by(sport, year)%>%
  dplyr::summarize(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop")%>%
  filter(sport %in% c("Swimming", "Athletics"))

ggplot(usa_fil, aes(x = year, y = total_winners, color = sport, group = sport)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Year",
    y = "Medal Count",
    title = "USA Olympic Medals Won in Athletics and Swimming",
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("#7B38EC", "#5CC0AB")) +
  scale_x_continuous(limits = c(1896, 2016), breaks = seq(1896, 2016, by = 8)) +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260"),
    axis.title.y = element_markdown(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.title = element_blank(),
    legend.position = c(0.8, 0.18),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm")
  )
```
---
class: middle, center

```{r wrangle-gdp, echo = FALSE, warning = FALSE}
detach(package:plyr, unload = TRUE)
gdp <- read_csv("../data/gdp-per-capita-maddison-2020.csv", show_col_types = FALSE)

gdp %<>%
  janitor::clean_names() %>%
  mutate(entity = case_when(
    entity == "United States" ~ "USA",
    entity == "United Kingdom" ~ "UK",
    entity == "Congo" ~ "Republic of Congo",
    entity == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo",
    entity == "Cote d'Ivoire" ~ "Ivory Coast",
    entity == "Trinidad and Tobago" ~ "Trinidad",
    entity == "North Macedonia" ~ "Macedonia",
    entity == "Czechia" ~ "Czech Republic",
    TRUE ~ entity
  ))

olympics_gdp <- olympics %>%
  group_by(region, date, season) %>%
  dplyr::summarise(total_winners = sum(medal_winner), total_score = sum(medal_score), .groups = "drop") %>%
  rename(entity = region) %>%
  mutate(
    year = year(date),
    summer_highlight = ifelse(total_winners > 100, T, F),
    winter_highlight = ifelse(total_winners > 30, T, F)
  ) %>%
  left_join(gdp, by = c("entity", "year")) %>%
  rename(country = entity)
```

```{r faceted-gdp-plots-1, fig.asp = 1.2, echo = FALSE, out.width = "50%", warning = FALSE}
winter_highlight <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(2000:2020), total_winners > 30)

olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Winter", year %in% c(2000:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = winter_highlight, size = winter_highlight), alpha = .5, size = .75, show.legend = FALSE) +
  geom_text_repel(
    data = winter_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
    color = "#7B38EC", size = 3.5, family = "Oswald"
  ) +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(
    title = "Total Medals Won vs. GDP per capita",
    subtitle = "Winter Olympics, 2002-2014",
    y = "Total Medals Won",
    x = "GDP per capita (USD)",
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 50)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")
  ) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    legend.title = element_blank(),
    legend.position = c(0.8, 0.68),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
    axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14)
  )

summer_highlight <- olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Summer", year %in% c(2004:2020), total_winners > 100)
```
---
class: middle, center
```{r faceted-gdp-plots-2, fig.asp = 1.2, echo = FALSE, out.width = "50%", warning = FALSE}
olympics_gdp %>%
  filter(!is.na(gdp_per_capita), season == "Summer", year %in% c(2004:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = summer_highlight, size = summer_highlight),
    alpha = .5, size = .75, show.legend = FALSE
  ) +
  geom_text_repel(
    data = summer_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
    color = "#7B38EC", size = 3.5, family = "Oswald"
  ) +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(
    title = "Total Medals Won vs. GDP per capita",
    subtitle = "Summer Olympics, 2004-2016",
    y = "Total Medals Won",
    x = "GDP per capita (USD)",
    caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com"
  ) +
  scale_y_continuous(breaks = seq(from = 0, to = 300, by = 100)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")
  ) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    legend.title = element_blank(),
    legend.position = c(0.8, 0.68),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
    axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14)
  )
```
---
class: middle, center
```{r faceted-gdp-plots-3, fig.asp = 1.2, echo = FALSE, out.width = "50%", warning = FALSE}
olympics_gdp %>%
  filter(!is.na(gdp_per_capita), country != "Qatar", season == "Summer", year %in% c(2004:2020)) %>%
  ggplot(aes(x = gdp_per_capita, y = total_winners)) +
  geom_point(aes(color = summer_highlight, size = summer_highlight),
    alpha = .5, size = .75, show.legend = FALSE
  ) +
  geom_text_repel(
    data = summer_highlight, aes(x = gdp_per_capita, y = total_winners, label = country),
    color = "#7B38EC", size = 3.5, family = "Oswald"
  ) +
  facet_wrap(~year, ncol = 1) +
  scale_color_manual(values = c("#5CC0AB", "#7B38EC")) +
  labs(title = "Total Medals Won vs. GDP per capita",
       subtitle = "Summer Olympics, 2004-2016",
       y = "Total Medals Won",
       x = "GDP per capita (USD)",
       caption = "Source: Sports Reference & OlympStats\nCompiled by kaggle.com") +
  scale_y_continuous(breaks = seq(from = 0, to = 300, by = 100)) +
  scale_x_continuous(
    breaks = seq(from = 0, to = 90000, by = 30000),
    labels = label_dollar(scale = .001, prefix = "$", suffix = "K")
  ) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing.y = unit(0.4, "cm"),
    legend.title = element_blank(),
    legend.position = c(0.8, 0.68),
    legend.background = element_rect(size = 0.3),
    legend.margin = margin(1, 5, 5, 5),
    legend.key.size = unit(0.5, "cm"),
    plot.title = element_text(family = "Oswald", color = "#092260", size = 20, hjust = 0.5),
    plot.caption = element_text(family = "Oswald", color = "#092260", size = 11),
    strip.text = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    plot.subtitle = element_text(family = "Oswald", color = "#092260", size = 14, hjust = 0.5),
    axis.text.x = element_text(family = "Oswald", color = "#092260"),
    axis.title.x = element_markdown(family = "Oswald", color = "#092260", size = 14),
    axis.text.y = element_text(family = "Oswald", color = "#092260", size = 8),
    axis.title.y = element_text(family = "Oswald", color = "#092260", size = 14),
    legend.text = element_text(family = "Oswald", color = "#092260", size = 14)
  )
```


